---
title: "Library Prep. Report"
date: "`r Sys.Date()`"
output: html_document
params:
  data: NA
---

```{r echo = FALSE, message = FALSE}
library(ggplot2)
library(knitr)
library(dplyr)

df <- params$data
```

:::: {style="display: flex; align-items: center;"}

::: {style="flex-basis: 200px; flex-grow: 2"}
```{r standards_plot, echo = FALSE, warning = FALSE, fig.height=3, out.width="100%"}

#                            12.7 <- dilution_lines (geom_text)
#      sample summary  _________________
#             v        |      ^         | <- vert_lines (geom_segment)
#       O .. ..  .     O      |         O <- standard_summary
#       |______________|      |------------- dilution_lines
#             10.3

dilution_lines <- df$standard_summary |>
  dplyr::filter(.data$name == "quant_actual") |>
  dplyr::mutate(line_start = 1/.data$value,
                line_end = dplyr::lag(.data$line_start),
                dil = dplyr::lag(.data$dil),
                y = rep_len(c(1.1, 0.9), 5),
                y_text = rep_len(c(1.15, 0.85), 5)) |>
  dplyr::filter(!is.na(.data$line_end)) |>
  dplyr::rowwise() |>
  dplyr::mutate(mid = sqrt(.data$line_start * .data$line_end))

vert_lines <-
  dplyr::tibble(x = c(dilution_lines$line_start, dilution_lines$line_end)) |>
  dplyr::arrange(.data$x) |>
  dplyr::mutate(y = rep(c(1.1, 0.9, 1.1, 0.9), each = 2),
                yend = rep(c(1.05, 0.95, 1.05, 0.95), each = 2))

df$standard_summary |>
  ggplot2::ggplot(aes(x = 1/.data$value, y = 1, color = .data$name)) +
  ggplot2::geom_point(size = 10, alpha = 0.7) +
  ggplot2::scale_color_manual(values = c("#00AAAA", "#222222")) +
  ggplot2::geom_segment(data = dilution_lines,
                        aes(x = .data$line_start, xend = .data$line_end,
                            y = .data$y, yend = .data$y),
                        size = 1) +
  ggplot2::geom_segment(data = vert_lines,
                        aes(x = .data$x, xend = .data$x,
                            y = .data$y, yend = .data$yend),
                        color = "#00AAAA", size = 1) +
  ggplot2::geom_text(data = dilution_lines,
                     aes(label = round(.data$dil, 1),
                         y = .data$y_text,
                         x = .data$mid),
                     size = 8) +
  ggplot2::scale_x_log10() +
  ggplot2::theme_minimal() +
  ggplot2::theme(panel.grid = element_blank(), 
                 axis.title = element_blank(), 
                 axis.text = element_blank(), 
                 legend.position = "none") +
  ggplot2::coord_cartesian(ylim = c(0.8, 1.2)) +
  ggplot2::geom_point(data = df$sample_summary, aes(x = 1/.data$quantity_mean, y = 1), color = "red")


```
:::


::: {style="flex-basis: 100px; flex-grow: 1"}
Theoretical (perfectly diluted) samples are in gray, true dilutions are in blue. Samples are in red. Dilutions between standards shown in text.
:::

::::


:::: {style="display: flex; align-items: center;"}

::: {style="flex-basis: 200px; flex-grow: 2"}
```{r outliers_plot, echo = FALSE, warning = FALSE, fig.height=5, fig.width = 7.5, out.width="100%"}
df$outliers |>
  dplyr::group_by(.data$sample_name) |>
  dplyr::mutate(keep_logi = !is.na(.data$keep),
                z = (.data$ct-.data$adj_mean)/.data$adj_sd,
                overflow = abs(.data$z) > 10,
                z_plot = dplyr::case_when(.data$z > 10 ~ 10,
                                          .data$z < -10 ~ -10,
                                          TRUE ~ .data$z),
                label = dplyr::case_when(.data$z > 10 ~ paste(as.character(round(.data$z, 0)),">>>"),
                                         .data$z < -10 ~ paste("<<<", as.character(round(.data$z, 0))),
                                         TRUE ~ NA_character_)) |>
  dplyr::filter(!is.na(.data$sample_name)) |> 
  ggplot2::ggplot(aes(x = .data$z_plot, y = .data$sample_name, color = .data$keep_logi, shape = .data$overflow)) +
  ggplot2::scale_color_manual(values = c("#666666", "#00AAAA")) +
  ggplot2::scale_shape_manual(values = c(16, NA)) +
  ggplot2::geom_text(aes(label = .data$label), size = 5) +
  ggplot2::geom_rect(xmin = -3, xmax = 3, ymin = 0, ymax = nrow(df$outliers), fill = "#AACCCC",  alpha = 0.1,  color = NA) +
  ggplot2::geom_vline(xintercept = 0, color = "#555555") +
  ggplot2::geom_point(size = 5) +
  ggplot2::xlab("Z-score") +
  ggplot2::coord_cartesian(xlim = c(-11.5, 11.5), 
                           ylim = c(0.5, length(unique(df$outliers$sample_name)) + 0.5), 
                           expand = FALSE) +
  ggplot2::theme(panel.grid.major.x = element_blank(), 
                 axis.title.y = element_blank(), 
                 panel.grid.minor.x = element_blank(), 
                 legend.position = "none")
```
:::

::: {style="flex-basis: 100px; flex-grow: 1"}
CT-values as represented by Z-score. Blue dots represent samples not considered outliers. Shaded blue region is +/- three standard deviations, which is considered the cutoff for outliers. Black dots are samples considered outliers. Far outliers (greater than 11 standard deviations) are represented by arrows (>>>) with the true value displayed beside it. Z scores are calculated after removing the outlier, if any. 
:::

::::


:::: {style="display: flex; align-items: center;"}

::: {style="flex-basis: 200px; flex-grow: 2"}
```{r concentration_plot, echo = FALSE, warning = FALSE, out.width="100%"}
df$sample_summary |>
  ggplot2::ggplot(aes(x = .data$sample_name, y = .data$concentration_mean)) +
  ggplot2::geom_point(color = "#00AAAA", size = 5) +
  ggplot2::theme(axis.title = element_blank(), 
                 legend.position = "none", 
                 axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5))

```
:::

::: {style="flex-basis: 100px; flex-grow: 1"}
Library concentrations across samples
:::

::::

:::: {style="display: flex; align-items: center;"}

::: {style="flex-basis: 200px; flex-grow: 2"}
```{r slope_plot, echo = FALSE, warning = FALSE, message = FALSE, out.width="100%"}

slope_text <-
  dplyr::tibble(x = -2.5, 
                y = 20,
                label = paste(paste("Slope:", round(df$standards$slope[1], 2)),
                              paste("R\u00B2:", round(df$standards$r2[1], 2)),
                              paste0("Efficiency: ", round(df$standards$efficiency[1], 1), "%"),
                              sep = "\n"))

df$standards |>
  ggplot2::ggplot(aes(x = log10(.data$quantity), y = .data$ct)) +
  ggplot2::geom_point(color = "#00AAAA", size = 5) +
  ggplot2::geom_smooth(method = "lm", se = F, color = "#666666", formula = "y ~ x") +
  ggplot2::geom_text(data = slope_text, aes(x = .data$x,  y = .data$y, label = .data$label), size = 8) +
  ggplot2::xlab("Log10(Quantity)") +
  ggplot2::ylab("CT")
```
:::

::: {style="flex-basis: 100px; flex-grow: 1"}
Log of library quantities vs CT values. An optimal slope is -3.32. An optimal efficiency is 100%. Acceptable efficiencies are >80%.
:::

::::

```{r sample_summary, echo = FALSE}
df$sample_summary |> 
  dplyr::select(-"quantity_mean") |>
  dplyr::rename("Sample Name" = .data$sample_name,
                "Concentration" = .data$concentration_mean) |> 
  knitr::kable()
```


